Spatiotemporal models of `r paste(params$species)` biomass with time-varying depth and trawl footprint

### 2020-01-01 analysis/VOCC/2-biomass-depth-only-model.Rmd
revise rmd for depth-only model to plot residuals, save predictions, and save time-varying depth plots

---
title: "Spatiotemporal models of `r paste(params$species)` biomass with time-varying depth and trawl footprint"
author: "Philina English"
date: "`r Sys.Date()`"
output: html_document
params: 
    species: "Redbanded Rockfish"
    region: "All synoptic surveys"
    knots: 600
    covariates: "+trawled+as.factor(ssid) + trawled:as.factor(ssid)" # additional fixed-effect covariates
    covs: "-trawled-ssid-reml" # covariate string used to name model rds 
    update_model: TRUE
---

@@

  adult_biomass <- sdmTMB::sdmTMB(data=data,
    adult_formula,
    time_varying = ~ 0 + depth_scaled + depth_scaled2, #+ trawled
    time = "year", spde = spde,
    family = tweedie(link = "log"),
    #ar1_fields = TRUE,
    include_spatial = TRUE,
    reml = TRUE,
    enable_priors = priors,
    #control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
    silent = FALSE
  )

@@

if (any(names(data) == "adult_density")) {
#if (length(unique(fish$maturity_code)) > 2) {
  nd_all <- readRDS(paste0("data/nd_all_synoptic.rds"))
  adult_biomass <- readRDS(paste0("data/", spp,
    "/mod-mat-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"))

@@

  adult_predictions <- predict(adult_biomass, newdata = nd)
  saveRDS(adult_predictions, file = paste0("data/", spp,
    "/predictions-", spp, covs, "-", ssid_string, "-mature-biomass-prior-", priors, ".rds"))

  imm_biomass <- readRDS(paste0("data/", spp,
    "/mod-imm-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
  ))

@@

  imm_predictions <- predict(imm_biomass, newdata = nd)
  saveRDS(imm_predictions, file = paste0("data/", spp,
    "/predictions-", spp, covs, "-", ssid_string, "-imm-biomass-prior-", priors, ".rds"))

@@

>>>>

---
title: "Spatiotemporal models of `r paste(params$species)` biomass with time-varying depth and trawl footprint"
author: "Philina English"
date: "`r Sys.Date()`"
output: html_document
params: 
    #species: "Redbanded Rockfish"
    #species: "Pacific Ocean Perch"
    #species: "Pacific Cod"
    #species: "Petrale Sole"
    species: "Pacific Halibut"
    region: "All synoptic surveys"
    knots: 500
    covariates: "" # +as.factor(ssid)
    #additional fixed-effect covariates
    #covs: "-ssid-only" # covariate string used to name model rds 
    covs: "-tv-depth-only"
    #covs: "-trawled-ssid"
    update_model: FALSE
---

@@

  adult_biomass <- sdmTMB::sdmTMB(data=data,
    adult_formula,
    time_varying = ~ 0 + depth_scaled + depth_scaled2, #+ trawled
    time = "year", spde = spde,
    family = tweedie(link = "log"),
    #ar1_fields = TRUE,
    include_spatial = TRUE,
    #reml = TRUE,
    enable_priors = priors,
    #control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
    silent = FALSE
  )

@@

if (any(names(data) == "adult_density")) {
#if (length(unique(fish$maturity_code)) > 2) {
  nd_all <- readRDS(paste0("data/nd_all_synoptic.rds"))

  try({adult_biomass <- readRDS(paste0("data/", spp,
    "/mod-mat-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"))

@@

  adult_predictions <- predict(adult_biomass, newdata = nd)
  saveRDS(adult_predictions, file = paste0("data/", spp,
    "/predictions-", spp, covs, "-", ssid_string, "-mature-biomass-prior-", priors, ".rds"))
  })

  try({
    imm_biomass <- readRDS(paste0("data/", spp,
    "/mod-imm-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"))

@@

  imm_predictions <- predict(imm_biomass, newdata = nd)
  saveRDS(imm_predictions, file = paste0("data/", spp,
    "/predictions-", spp, covs, "-", ssid_string, "-imm-biomass-prior-", priors, ".rds"))
 })

@@

```{r}
  try({ biomass <- readRDS(paste0("data/", spp,
    "/mod-mat-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"))
  })

  try({ biomass <- readRDS(paste0("data/", spp,
    "/model-total-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"))
  })

point_predictions <- predict(biomass)
point_predictions$residuals <- residuals(biomass)

saveRDS(point_predictions, file = paste0("data/", spp,
    "/check-mod-predictions-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"))

```

```{r}
depth_only_predictions <- readRDS(paste0("data/", spp,
    "/check-mod-predictions-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"))

depth_only_predictions <- filter(depth_only_predictions, 
  #ssid %in% c(1,3)) %>% filter (
  year > 2007)

 try({ 
   g <- ggplot(depth_only_predictions, aes(adult_density, residuals, colour = adult_density)) + 
          geom_point() + scale_x_continuous(trans = 'log10') + facet_wrap(~year)
 })

 try({ 
   g <- ggplot(depth_only_predictions, aes(density, residuals, colour = density)) + 
          geom_point() + scale_x_continuous(trans = 'log10') + facet_wrap(~year)
 })

g
```

```{r}
 try({ 
   g2 <- ggplot(depth_only_predictions, aes((est), residuals, colour = adult_density)) +
            geom_point() + #scale_x_continuous(trans = 'log10') + 
            facet_wrap(~year)
 })

 try({ 
   g2 <- ggplot(depth_only_predictions, aes((est), residuals, colour = density)) +
            geom_point() + #scale_x_continuous(trans = 'log10') + 
            facet_wrap(~year)
 })

g2
```

```{r}
ggplot(depth_only_predictions, aes(X, Y, colour = (residuals))) + 
  geom_point() + 
  scale_colour_gradient2() + 
  scale_x_continuous(trans = 'log10') + 
  facet_wrap(~year) 
```

### Build time-varying depth plots
#### Model without environmental covariates
```{r }
max_depth_found <- max(data[data$present==1,]$raw_depth, na.rm = TRUE)
rm(depth_model_list)
rm(adult_biomass)
rm(imm_biomass)
rm(total_biomass)

try({  adult_biomass <- readRDS(paste0(
    "data/", spp,
    "/mod-mat-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
  ))})  
#adult_biomass<-sdmTMB:::update_model(adult_biomass)

  try({  imm_biomass <- readRDS(paste0(
    "data/", spp,
    "/mod-imm-biomass-", spp, covs, "-1n3n4n16-prior-FALSE.rds"
  ))
  #imm_biomass<-sdmTMB:::update_model(imm_biomass) 
 # depth_model_list <- list(adult = adult_biomass, imm = imm_biomass)
})

    if (!exists("adult_biomass")) {
      try({
        total_biomass <- readRDS(paste0(
    "data/", spp,
    "/model-total-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
  ))
      })
    }

if (!exists("adult_biomass")) adult_biomass <- NULL
if (!exists("imm_biomass")) imm_biomass <- NULL
if (!exists("total_biomass")) total_biomass <- NULL

  depth_model_list <- list(adult = adult_biomass, imm = imm_biomass, total = total_biomass)
  depth_model_list <- depth_model_list[!sapply(depth_model_list, is.null)]

d <- list()
depth_plots <- list()
for (i in seq_len(length(depth_model_list))) {
 if(depth_model_list[[i]]$model$convergence==0){
  d[[i]] <- time_varying_density(depth_model_list[[i]], predictor = "depth")

  if (length(d[[i]]) == 0) {
    depth_plots[[i]] <- grid::grid.rect(gp = grid::gpar(col = "white"))
  } else {
    d[[i]]$x <- exp(d[[i]]$x)
    depth_plots[[i]] <- plot_mountains(d[[i]], variable_label = "Depth (without environmental variables)", xlimits = c(0, max_depth_found)) +
      ggtitle(paste(species, names(depth_model_list[i])))
  }
 } else {
    depth_plots[[i]] <- grid::grid.rect(gp = grid::gpar(col = "white"))
  } 
}
print(depth_plots)
```

```{r}
png(
  file = paste0(
    "figs/", spp,
    "/depth-", spp, covs, ".png"
  ),
  res = 600,
  units = "in",
  width = 8.5,
  height = 6
)
gridExtra::grid.arrange(
  grobs = c(depth_plots),
  nrow = 2,
  top = grid::textGrob(paste(species, "(", covs, ")"))
)
dev.off()
```

```{r}
depth_model_list
```
